Ugly DECOY
==========

## Access to the Babel Clinical Data Warehouse

```{r babel.access,echo=FALSE}
# see install_RPostgres.Rmd for bleeding-edge details
library(DBI)
library(RPostgres)

# based on DBI::getQuery and RPostgres::dbSendQuery
# https://github.com/rstats-db/DBI/blob/7a0ad76dea21a846cee62f67108ab8e8b7d60a49/R/DBConnection.R#L133
makePreparedQuery <- function(conn) {
  function(statement, ...) {
    params = list(...)

    rs <- dbSendQuery(conn, statement)
    on.exit(dbClearResult(rs))
    
    if (length(params) > 0) {
      dbBind(rs, params)
      }
    
    df <- dbFetch(rs, n = -1, ...)
    if (!dbHasCompleted(rs)) {
      warning("Pending rows", call. = FALSE)
      }
    
    df
    }
  }


babel.access <- function(myCredsFile) {
  creds <- source(myCredsFile)$value
  function() {
    dbConnect(RPostgres::Postgres(), dbname="i2b2", host="localhost",
              user=creds$user, password=creds$password, port=creds$port)
    }
  }
```

Create a `babel_creds.R` file a la:

```
list(user="me", password="sekret", port=5000)
```

where 5000 is the local end of an ssh tunnel to babel's postgres port (5432).

Then we can get a connection and run SQL queries:

```{r cdw}
myAccess = babel.access(myCredsFile='babel_creds.R')
cdw = myAccess()
dbGetQuery(cdw, "select 1+1")

cdwq <- makePreparedQuery(cdw)
cdwq("select 2 + 2")
cdwq("select $1 + 5", p1=3)
```

The indirection thru `myAccess` encapsulates your credentials from your
exploratory computing session.

## Basic Demographics

There are quite a few metadata tables:

```{r}
table_access <- cdwq(
  "select distinct c_name, c_table_name, c_fullname, c_totalnum
  from i2b2metadata.table_access")

nrow(table_access)
```

Let's pick out the ones relevant to demographics:

```{r}
dem_tables <- table_access[grep('demog', table_access$c_name, ignore.case=TRUE), ]
head(dem_tables)
```

Next, we summarize age distribution:

```{r summary_nominal}

summary_nominal <- function(table, path_pattern) {
  cdwq(
    sub('&&TABLE', table,
        "select coalesce(sum(c_totalnum), 0) pat_qty, c_fullname, c_basecode
        from i2b2metadata.&&TABLE meta
        where meta.c_fullname like ('%' || $1 || '%') escape '$'
        and length(c_basecode) > 0
        group by c_fullname, c_basecode"),
    p1=path_pattern)
  }

age_dist <- summary_nominal(dem_tables[grep('KUMC', dem_tables$c_name),]$c_table_name, 'Demographics\\Age\\_')
age_dist$age <- as.numeric(sub('DEM|AGE:', '', age_dist$c_basecode, fixed=TRUE))
head(age_dist)
```

*something odd happens above 88 yrs old; let's leave that aside for the plot...*

```{r age.plot}
with(subset(age_dist, age < 88),
     plot(age, pat_qty))
```


## Synthesize patients

```{r synpat}

synpat <- function(qty, age_dist) {
  pat <- data.frame(patient_num=1:qty)
  pat$age <- with(age_dist, sample(age, qty, prob=pat_qty))
  pat
  }

spat <- synpat(100, age_dist)
head(spat)
```

## Save patients in CSV format

```{r write_patients}
write.csv(spat, file="patient.csv", row.names=FALSE)
```
